---
title: "PopDemo"
author: "Maria Berg"
date: "2025-03-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)


# Filter out the year 2020
df <- HousingMarket%>% filter(Year != 2020) 

# Keep only one row per year per region (e.g., taking the first row for each year)
df_yearly <- df %>%
  group_by(Region, Year) %>%
  summarise(across(starts_with("PopPct"), mean, na.rm = TRUE), .groups = "drop")  # Use mean to aggregate



# Reshape the data for stacked bar chart
df_long <- df_yearly %>%
  pivot_longer(cols = starts_with("PopPct"), 
               names_to = "Demographic", 
               values_to = "Percentage") 

# Clean column names
df_long$Demographic <- gsub("PopPct_", "", df_long$Demographic) 
df_long$Demographic <- gsub("\\.", " ", df_long$Demographic) 


df_long <- df_long %>%
  filter(!Demographic %in% c("TOTAL POPULATION", "18 YEARS AND OVER"))  # Exclude these demographics

df_long <- df_long %>%
  mutate(Region = factor(Region, levels = c("Philadelphia", "Chicago", "Los Angeles", "Boston", "Seattle")))

# Create the faceted stacked bar chart
ggplot(df_long, aes(x = factor(Year), y = Percentage, fill = Demographic)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ Region, nrow = 2, scales = "free_x") +
  #facet_wrap(~Region) +
  scale_fill_viridis_d(option = "C", begin = 0.1, end = 0.9) +  # Colorblind-friendly palette
  labs(title = "Demographic Distribution by Year for Each Region",
       x = "Year",
       y = "Percentage of Population",
       fill = "Demographic") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    #legend.position = "bottom",
    legend.position = c(0.999, 0.08),  # Adjust x and y to move legend into empty space
    legend.justification = c(1, 0),
    legend.text = element_text(size = 9),  # Adjust legend text size
    legend.title = element_text(size = 13),  # Adjust legend title size
    strip.placement = "outside",  # Ensures proper spacing of facet labels
    strip.text = element_text(size = 12),  # Adjust facet label size if needed
    axis.line = element_line(size = 0.5, color = "black")  # Increase thickness of axis lines
  ) 


```

The graph presents the demographic distribution of five major U.S. cities—Philadelphia, Chicago, Los Angeles, Boston, and Seattle—over multiple years. Each subplot represents a city, with stacked bars to illustrating the proportion of different racial and ethnic groups in the population each year. 


The visualization helps to observe long-term demographic trends, shifts in diversity, and changes in racial composition across these metropolitan areas.




```{r}
library(ggplot2)
library(dplyr)
library(tidyr)


# Filter out the year 2020
df <- HousingMarket%>% filter(Year != 2020) 

# Keep only one row per year per region (e.g., taking the first row for each year)
df_yearly <- df %>%
  group_by(Region, Year) %>%
  summarise(across(starts_with("PopEst"), mean, na.rm = TRUE), .groups = "drop")  # Use mean to aggregate



# Reshape the data for stacked bar chart
df_long <- df_yearly %>%
  pivot_longer(cols = starts_with("PopEst"), 
               names_to = "Demographic", 
               values_to = "Population") 

# Clean column names
df_long$Demographic <- gsub("PopEst_", "", df_long$Demographic) 
df_long$Demographic <- gsub("\\.", " ", df_long$Demographic) 


df_long <- df_long %>%
  filter(!Demographic %in% c("TOTAL POPULATION", "18 YEARS AND OVER"))  # Exclude these demographics

df_long <- df_long %>%
  mutate(Region = factor(Region, levels = c("Philadelphia", "Chicago", "Los Angeles", "Boston", "Seattle")))

# Create the faceted stacked bar chart
ggplot(df_long, aes(x = factor(Year), y = Population, fill = Demographic)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ Region, nrow = 2, scales = "free_x") +
  #facet_wrap(~Region) +
  scale_fill_viridis_d(option = "C", begin = 0.1, end = 0.9) +  # Colorblind-friendly palette
  labs(title = "Demographic Distribution by Year for Each Region",
       x = "Year",
       y = "Percentage of Population",
       fill = "Demographic") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    #legend.position = "bottom",
    legend.position = c(0.999, 0.08),  # Adjust x and y to move legend into empty space
    legend.justification = c(1, 0),
    legend.text = element_text(size = 9),  # Adjust legend text size
    legend.title = element_text(size = 13),  # Adjust legend title size
    strip.placement = "outside",  # Ensures proper spacing of facet labels
    strip.text = element_text(size = 12),  # Adjust facet label size if needed
    axis.line = element_line(size = 0.5, color = "black"),  # Increase thickness of axis lines
  ) 
```

